---
title       : Definite Integral Approximation
subtitle    : An Application for the Developing Data Products Course Project
author      : Chris Breaux
job         : 
framework   : io2012      # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Introduction

Approximating Definite Integrals is an application that illustrates the concept of integration from basic calculus. The following is the basic workflow:

1. The user inputs a function
2. A plot of the corresponding curve is shown
3. The area under the curve is estimated by drawing a user-specified number of rectangles

---

## Example Graph Output
Consider the following example:

Function = $x^2/e^x$, Number of Rectangles = 10

```{r echo=FALSE,fig.align='center'}

expr = parse(text="x^2/exp(x)")
fn = function(x) eval(expr)
ll = 0
ul = 10
n = 10

fnx = seq(from = ll, to = ul, length.out = n+1)
fnx = fnx[-length(fnx)]
fny = fn(fnx)

curve(fn, from = ll, to = ul)
rect(fnx,0,fnx+(ul-ll)/n,fny,border=TRUE)
```



---

## Better Approximations to the Definite Integral

When we increase the number of rectangles in the plot, the sum of the areas of the rectangles becomes a closer approximation of the the true value of the definite integral. 

Function = $x^2/e^x$, Number of Rectangles = 100
```{r echo=FALSE,fig.align='center'}
n = 100
fnx = seq(from = ll, to = ul, length.out = n+1)
fnx = fnx[-length(fnx)]
fny = fn(fnx)

curve(fn, from = ll, to = ul)
rect(fnx,0,fnx+(ul-ll)/n,fny,border=TRUE)
```

---

## Other options

The application also outputs the sum of the areas of the rectangles and the true value of the definite integral for comparison.

Also, you are able to change the upper and lower bounds over which the integral is evaluated.

However, the most powerful use of the application is that you are free to use an arbitrary arithmatic expression in R to be evaluated. For instance, the following functions are all valid:

* x ^ 2 / exp(x)

* 1 + x + x ^ 2 / 2 + x ^ 3 / 6

* ifelse(x > 3, 3, x)
